Growth and inequality in public good provision \textbar{} an extended replication

### 2022-10-09 analysis/quarto/paper.tex
write results:replication

\begin{keyword}
    Replication study \sep Non-convenience sample \sep Open
science \sep Dynamic public good game \sep Online experiment \sep 
    Generalizability
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, sharp corners, enhanced]}{\end{tcolorbox}}\fi

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\textbf{Hintergrund:} Ich habe diesen
\href{https://groups.google.com/g/esa-announce/c/-_2OmbkdxEk}{post}
gesehen und überlege meine Replikation entsprechend
\href{https://www.sciencedirect.com/journal/journal-of-behavioral-and-experimental-economics/about/call-for-papers\#transparency-reproducibility-and-generalizability-of-behavioral-economics-experiments}{hier}
einzureichen. Ist das realistisch? Wenn ja, wie soll man die Story
aufziehen? Soll man denen mal schreiben und fragen, ob noch Platz ist?

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

@@

\hypertarget{pre-registered-gmtv-replication}{%
\section{Pre-registered GMTV
Replication}\label{pre-registered-gmtv-replication}}

\hypertarget{provision-of-the-public-good}{%
\subsection{Provision of the public
good}\label{provision-of-the-public-good}}

First, we ask whether the samples differ with respect to their initial
contributions to the public good. Is our replication sample more
pro-social than the original sample? Figure 1 reveals, that it is not.
The distributions of both samples look fairly similar. Both samples
contributed 10 tokens, that is, 50\% of their endowments on average
(median and mean).\footnote{The two-sided rank sum test (comparing
  differences between samples) yields a p-Value of 0.3926 for the mean
  contribution in first round of the game.} Moreover, both samples'
initial contributions resemble initial contributions participants
usually make in the standard game with partner matching.\footnote{See
  Figure 3B in \citet{fehrgaechter2000} (p.989), for instance.} In the
dynamic game presented here, we are particularly interested in the
subsequent periods because differences add up exponentially. Do the two
groups remain similar?

\includegraphics{paper_files/figure-pdf/firstRoundViz-1.pdf}

In particular, do the two samples' contributions follow the same pattern
over time? The answer is \emph{no}. Figure 2 illustrates that the
samples make similar contributions at the beginning and the end of the
game but behave differently in between. More precisely, the left
panel--depicting the average contributions in absolute terms--shows that
the original sample contributed substantially more than the replication
sample \emph{in all but the first and last rounds}. For this reason, the
original sample's behavior differs from the replication sample's
behavior in two aspects: it contributes more and exhibits a considerable
drop in the last period (whereas the replication sample's contributions
flatten). Note that increasing contributions over time imply increasing
endowments over time. Hence, increasing absolute contributions do not
necessarily imply that participants contribute increasing shares of
their endowments. The right panel in Figure 1 shows the share of overall
endowments contributed over time where both samples exhibit a similar
pattern: they decline and do not stabilize. The original sample's
behavior differs from the replication sample's behavior in one aspect:
it declined slower (to the same level in the last period).

@@

Moreover, both samples' contributions resemble contributions
participants usually make in the standard game with partner
matching.\footnote{The right panel is thus, comparable to the
  visualizations \emph{and results} in the standard game. See, for
  instance, Figure 1B in \citet{fehrgaechter2000} (p.986).} In the
dynamic game presented here, different paths lead to different levels of
wealth -- even if they share the same start- and end-points. We are
thus, more interested in the implications contributions have for wealth
generation and growth. So, how do the samples compare with respect to
wealth/income/endowments/stock?

\hypertarget{wealth-creation}{%
\subsection{Wealth Creation}\label{wealth-creation}}

How do the different contribution-paths translate to wealth?\footnote{To
  measure growth, we define a variable called \emph{stock} which sums
  the endowments of all participants in a given group at the end of the
  round (that is, after the contributions have been made, multiplied and
  redistributed).} Given that the original sample contributed more in
most of the periods, one would expect the respective groups to be
considerably more wealthy. Figure 3 indicates that they tend to be
insignificantly\footnote{The two-sided rank sum test (comparing
  differences between samples) yields a p-Value of 0.1356 for the mean
  stock in last round of the game.} richer.

\includegraphics{paper_files/figure-pdf/stockDistributionViz-1.pdf}

As the grey lines indicate, an average group in the original sample
accumulated about 478 tokens. In contrast, an average group in the
replication sample accumulated about 380 tokens. However, none of the
groups were efficient: the maximal wealth that can be reached in round
10 under full cooperation is approximately 4613 tokens or 230 Euro per
group.

\includegraphics{paper_files/figure-pdf/growthHeterogeneityViz-1.pdf}

While there is clearly growth, groups do not realize the potential
efficiency as the replication groups reach on average a level of 379
tokens out of 4613 maximally possible or 8.2\%. As in the original data,
there is large heterogeneity with the richest group reaching 1425 tokens
whereas the poorest group ends up with 92 tokens.

We thus, observe growth even for the poorest group and spot
heterogeneity even within data sources.

Figure 2 shows the dynamics of wealth over time.\footnote{Unlike GMTV I
  present no median split but only the main results.}

@@

{\centering \includegraphics{paper_files/figure-pdf/plotStock-1.pdf}

@@

The Figure illustrates that (in both data sources) growth was continuous
and surprisingly linear, given the exponential character of the game's
design.

To sum up, our groups also tend to be poorer, and median wealth is
higher in GMTV. The difference in mean ranks is not significant
according to a two-sided ranksum test\footnote{The two-sided rank sum
  test (comparing differences between data sources) yields a p-Value of
  0.1356 for the mean wealth after the last round of the game.},
however.

To further assess the statistical significance of differences in means,
we run OLS regressions where we regress wealth on a treatment dummy for
\emph{Replication} (Table 3). These regressions show that differences in
means are only significant for below median groups. This implies that
the poor groups in our data are even less wealthy than GMTV's poor
groups.

\begin{table}[!htbp] \centering 
  \caption{} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{\textit{Dependent variable:}} \\ 
\cline{2-4} 
\\[-1.8ex] & \multicolumn{3}{c}{Wealth} \\ 
 & All & Below median & Above median \\ 
\hline \\[-1.8ex] 
 Replication & $-$98.26 & $-$59.41$^{***}$ & $-$138.21 \\ 
  & (101.21) & (18.32) & (166.67) \\ 
  & & & \\ 
 Constant & 478.09$^{***}$ & 234.70$^{***}$ & 731.00$^{***}$ \\ 
  & (75.58) & (13.99) & (124.73) \\ 
  & & & \\ 
\hline \\[-1.8ex] 
Observations & 52 & 24 & 25 \\ 
R$^{2}$ & 0.02 & 0.32 & 0.03 \\ 
Residual Std. Error & 362.49 & 44.24 & 413.67 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{3}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

As such, our results look familiar to GMTV's findings, where poor groups
grew ever so slightly, while only rich groups experienced growth (that
became exponential only after the 10th round).

>>>>

\begin{keyword}
    Replication study \sep Non-convenience sample \sep Open
science \sep Dynamic public good game \sep Online experiment \sep 
    Generalizability
\end{keyword}
\end{frontmatter}\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, breakable, sharp corners, enhanced, interior hidden]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

In the terminology of \citet{Hamermesh2007}, I ran both a \emph{pure} as
well as a \emph{scientific replication} of one treatment arm of
\citet{GMTV2017}'s dynamic public good game. The pure replication
re-analyzes the original data.
\protect\hyperlink{A:-Pure-Replication}{Appendix A} documents
\href{}{errors I identified} in the original paper. The scientific
replication, where I utilize a different sample drawn from a different
population in a different situation, is described in this section.

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

Participants entered the experiment at appointed times remotely from
home. They first saw a welcome screen. After agreeing to the privacy
policy, they could proceed to the instructions individually. Having read
these instructions, each participant has also seen a demo-screen
explaining the user interface. Before proceeding, they had to answer six
comprehension questions correctly. Subsequently, they saw a waiting
screen until they could be matched with three other participants, who
have answered the comprehension questions correctly. Once matched, they
were exposed to the decision screen over ten periods. At the end of the
last period, participants saw results of all periods. Subsequently, they
were exposed to a voluntary climate action, where they could donate
(some of) their earnings to offset carbon dioxide. Subsequently, I
elicited risk preferences \citep{HoltLaury2002} and finished with
\citet{GMTV2017}'s questionnaire.

\hypertarget{experimental-design}{%
\subsection{Experimental Design}\label{experimental-design}}

As in the NOPUNISH 10 Period treatment arm of \citet{GMTV2017}, I ran
sessions with groups of four (\(i \in I=\{1,2,3,4\}\)), an initial
endowment of \(N_i^1 = 20\) tokens, \(T=10\) rounds, a private account
with a return of \(1\) and a group account with a return of \(1.5\)
(\(\Rightarrow\) MPCR\(\equiv \frac{1.5}{4}\)), such that:

\[
N_i^{t+1}=N_i^t - c_i^t + \frac{1.5}{4}\sum_{j=1}^4 c_j^t
\] Instead of receiving fresh endowments every period, participants
receive one initial endowment only at the beginning of the first period.
A participant's subsequent endowment equals her profit from the current
period such that a decision in one period has consequences on her future
endowments. For this reason, the game is described as a \emph{dynamic}
public good game

\hypertarget{sample}{%
\subsection{Sample}\label{sample}}

I recruited the participants from the so called
\emph{\href{https://www.wiso.uni-hamburg.de/forschung/forschungslabor/umfragelabor/aktuelle-umfragen/hamburgpanel.html}{HamburgPanel}}
using HROOT \citep{hroot}. The panel is provided by the University of
Hamburg's Research Laboratory, which used a randomized last digits
approach to build the panel while drawing from the population of
citizens from Hamburg, Germany.

Describe sample properties here.

\hypertarget{software}{%
\subsection{Software}\label{software}}

The experiment was created using oTree \citep{oTree} and can be found on
\href{https://github.com/Howquez/coopUncertainty}{GitHub}.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{online-feasibility}{%
\subsection{Online Feasibility}\label{online-feasibility}}

How did the participants, who have never participated in a online group
experiment before, cope with the situation?

{\centering \includegraphics{paper_files/figure-pdf/plotTime-1.pdf}

\caption{Average Time Spent for each Contribution per Period}

\hypertarget{pre-registered-gmtv-replication}{%
\subsection{Pre-registered GMTV
Replication}\label{pre-registered-gmtv-replication}}

\textbf{Result 1.} \emph{The \texttt{NOPUNISH\ 10} treatment of
\citet{GMTV2017} can be replicated because the replication data resemble
the original data with respect to initial and final contributions,
wealth and growth as well as inequality between and within groups.}

This is remarkable given the different sample and language, the
different software and user interface as well as the online setting
during the COVID19 pandemic.

\hypertarget{contributions}{%
\subsubsection{Contributions}\label{contributions}}

First, I ask whether the samples differ with respect to their initial
contributions to the public good. Is our replication sample more
pro-social than the original sample? Figure 1 reveals that it is not.
The distributions of both samples look fairly similar. Both samples
contributed 10 tokens, that is, 50\% of their endowments on average
(median and mean).\footnote{The two-sided rank sum test (comparing
  differences between samples) yields a p-Value of 0.3926 for the mean
  contribution in first round of the game.} Moreover, both samples'
initial contributions resemble initial contributions participants
usually make in the standard game with partner matching.\footnote{See
  Figure 3B in \citet{fehrgaechter2000} (p.989), for instance.} However,
in the dynamic game presented here, we are particularly interested in
the subsequent periods because differences add up exponentially. Do the
two groups remain similar over the course of time?

{\centering \includegraphics{paper_files/figure-pdf/firstRoundViz-1.pdf}

\caption{Individual contributions to the dynamic public good in the
first period}

In particular, do the two samples' contributions follow the same path
over the 10 periods they played? The answer is \emph{no}. Figure 2
illustrates that the samples make similar contributions at the beginning
and the end of the game but behave differently in between. More
precisely, the left panel--depicting the average contributions in
absolute terms--shows that the original sample contributed substantially
more than the replication sample \emph{in all but the first and last
periods}. For this reason, the original sample's behavior differs from
the replication sample's behavior in two aspects: it contributes more
and exhibits a considerable drop in the last period (whereas the
replication sample's contributions flatten).

Note that increasing contributions over time imply increasing endowments
over time. Hence, absolute contributions do not us much about the
willingness to cooperate. For this reason, the right panel in Figure 2
shows the average \emph{share of endowments contributed} over time. Both
samples exhibit a similar pattern: they decline and do not stabilize.
However, both samples also differ with respect to one aspect: the
replication sample's share of contributions declines faster.

@@

Again, both samples' behavior resembles the contributions participants
usually make in the standard game with partner matching: contributions
equal approximately half of endowments in the very first period and
decrease to around 10\% of endowments by the last period.\footnote{The
  right panel is thus, comparable to the visualizations \emph{and
  results} in the standard game. See, for instance, Figure 1B in
  \citet{fehrgaechter2000} (p.986).} In the dynamic game presented here,
however, different paths lead to different levels of wealth -- even if
they share the same start- and end-points. I am thus, more interested in
the contributions' implications for wealth generation and growth.

\hypertarget{wealth-creation}{%
\subsubsection{Wealth Creation}\label{wealth-creation}}

How do the different contribution-paths translate into
wealth?\footnote{To measure wealth and growth, I define a variable
  called \emph{stock} which sums the endowments of all participants in a
  given group at the end of the round (that is, after the contributions
  have been made, multiplied and redistributed).} Given that the
original sample contributed more in most of the periods, one would
expect the respective groups to be considerably more wealthy. Figure 3
indicates just that. The grey lines show that an average group in the
original sample accumulated about 478 tokens. In contrast, an average
group in the replication sample accumulated about 380 tokens. This
difference is insignificant at conventional levels\footnote{The
  two-sided rank sum test (comparing differences between samples) yields
  a p-Value of 0.1356 for the mean stock in last round of the game.}
though.

@@

{\centering \includegraphics{paper_files/figure-pdf/stockDistributionViz-1.pdf}

\caption{Groups' income at the end of the game}

Although there clearly is growth, groups do not realize the maximal
potential efficiency: under full cooperation, a group can accumulate at
least 4613 tokens or EUR 230. This is depicted in the left panel of
Figure 4, where one can see the average wealth over time by sample. The
panel illustrates for both samples that growth was continuous and
surprisingly linear, given the exponential character of the game's
design. To sum up, the contribution behavior differed between samples.
In contrast, neither the eventual wealth nor the corresponding growth
paths differed. Differences in contribution behavior did, thus, not
translate in significantly different wealth outcomes.

Why? Perhaps because the heterogeneity within samples and across groups
has been too large to \emph{detect} a significant difference. The right
panel of Figure 4 depicts heterogeneity: In the replication sample, the
richest group earned 1425 tokens (which is about 1781\% of the initial
endowment) whereas the poorest group ends up with 92 tokens (115\%).
More broadly, the replication sample is characterized by inequality
between groups (\(SD_{Replication} =\) 336.06). The same holds true for
the original sample (\(SD_{Original} =\) 393.58). Hence, the
heterogeneity across groups does not differ between samples, which is
remarkable because the replication sample was drawn from a more
heterogeneous (non-convenience sample). Does it differ within groups?

{\centering \includegraphics{paper_files/figure-pdf/growthHeterogeneityViz-1.pdf}

\caption{Average wealth over time across samples.}

\hypertarget{inequality}{%
\subsubsection{Inequality}\label{inequality}}

Given the different samples and the possibility of endogenous
growth--which essentially is the main feature of the game--I ask whether
and how the inequality grows \emph{within} groups. Figure 5 illustrates
that inequality did grow: at the end of the game, the original and the
replication groups exhibit an average Gini coefficient of 0.23 and 0.22,
respectively.\footnote{The two-sided rank sum test (comparing
  differences between samples) yields a p-Value of 0.6059 for the mean
  Gini coefficient in last round of the game.} Because every participant
started with the same initial endowment (in \emph{Period 0}, so to
speak), every group started equally--with a Gini coefficient equaling
zero.

Figure 6 shows that this initial state of equality ended with the first
period already: both samples exhibit a stark incline in inequality
before the second period started. From then on, the respective Gini
coefficients grew slowly but continuously -- for both samples.

{\centering \includegraphics{paper_files/figure-pdf/giniDistributionViz-1.pdf}

\caption{Groups' Gini coefficients (within groups) at the end of the
game}

{\centering \includegraphics{paper_files/figure-pdf/unnamed-chunk-1-1.pdf}

\caption{Average Gini coefficient (within groups) over time across
samples}

\hypertarget{generalizability}{%
\subsection{Generalizability}\label{generalizability}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\hypertarget{a-pure-replication}{%
\subsection{A: Pure Replication}\label{a-pure-replication}}

\hypertarget{b-growth-and-inequality-exploratory}{%
\subsection{\texorpdfstring{B: Growth \emph{and} Inequality
(exploratory)}{B: Growth and Inequality (exploratory)}}\label{b-growth-and-inequality-exploratory}}

In contrast to \citet{GMTV2017}, I did not ask how rich and poor groups
differ. Instead, I was wondering, whether equal groups are wealthier.
More precisely, does the Gini coefficient correlate with growth and
wealth creation? To answer that question, Figure 7 applies a median
split showing equal and unequal groups' wealth over time.

{\centering \includegraphics{paper_files/figure-pdf/plotStockByGini-1.pdf}

@@
